Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-01-18 paper/paper.tex
explain RCE

\begin{figure}[h!]
  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
  \caption{Suricata rule for SYN Flood Attacks.}
  \label{fig:synflood-example}
\end{figure}

@@

\begin{figure}[h!]
  \vspace{1ex}
  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
  \caption{Suricata rule for Ping Scan.}
  \label{fig:pingscan-example}
  \vspace{1ex}  
\end{figure}

@@

\subsection{Remote Code Execution}
\label{sec:rce}
\label{sec:jboss}
\label{sec:content-example}

\Luc{"RCE refers to the process by which an agent can exploit a network vulnerability to run arbitrary code on a targeted machine or system. In an RCE attack, hackers intentionally exploit a remote code execution vulnerability to run malware. By prompting the targeted device to perform code execution, a hacker can run their own programming in its place. This programming can then enable them to gain full access, steal data, carry out a full distributed denial of service (DDoS) attack, destroy files and infrastructure, or engage in illegal activity. And as the term remote execution suggests, an RCE cyberattack can take place from any geophysical location."\\ Texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution} \\

\Mar{Note que a gente organizou estas secoes por tipo de ataque. Este
  titulo eh muito especifico.}
\Luc{JBoss é um servidor de aplicação Java muito popular, de codigo aberto. Esse ataque tem como alvo esses servidores e permite que seja feita execução remota de comandos. Os pacotes que estamos utilizando para testar esse ataque foram gerados pelo Nikto, mas o Metasploit tambem possui um modulo para ele. A regra de ouro para esse ataque é:

\begin{figure}[H]
  \lstinputlisting[language=C,numbers=none,keywords={content}]{adaptor-golden-rule.suricata}
  \label{fig:rule-example}
\end{figure}

A opção flow:established,to\_server; define que o match só será feito em pacotes que fazem parte de uma conexão já estabelecida entre um cliente e um servidor. Além disso, define também que os pacotes devem estar direcionados ao servidor, não ao cliente. Nos contents os modificadores nocase e http\_uri indicam que a detecção da string não será case sensitive e que essa string deverá se encontrar no campo de request uri.}

\Luc{... sobre comportamento do syrius ...}

\Fix{...please elaborate. We need on example that shows that the
  search can be non-trivial. In the examples above, we had to discard
  only one rule option!}

>>>>

\begin{figure}[h!]
  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
  \caption{Suricata rule for SYN Flood attack.}
  \label{fig:synflood-example}
\end{figure}

@@

\begin{figure}[h!]
  \vspace{1ex}
  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
  \caption{Suricata rule for Ping Scan attack.}
  \label{fig:pingscan-example}
  \vspace{1ex}  
\end{figure}

@@

\subsection{Remote Code Execution (RCE)}
\label{sec:rce}
\label{sec:jboss}
\label{sec:content-example}

%% Adaptacao do texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution

RCE refers to the family of attacks where hackers exploit a network
vulnerability to run arbitrary code---that could damage the system or
steal sensitive information---on a remote machine. JBoss~\cite{jboss}
is popular open source Java application server. This attacks enables
remote execution of code on the JBoss application server.
\Mar{@Lucas, por favor, observe o padrao usado nos exemplos
  anteriores. eh preciso explicar em alto nivel o ataque antes de
  explicar a regra!}

%% In an RCE attack, hackers
%% intentionally exploit a vulnerability in the system to run malware
%% By prompting the targeted device to perform code execution, a hacker
%% can run arbritrary code to
%% their own programming in its
%% place. This programming can then enable them to gain full access,
%% steal data, carry out a full distributed denial of service (DDoS)
%% attack, destroy files and infrastructure, or engage in illegal
%% activity. And as the term remote execution suggests, an RCE
%% cyberattack can take place from any geophysical location."\\

\begin{figure}[H]
  \lstinputlisting[language=C,numbers=none,keywords={content,flow}]{adaptor-golden-rule.suricata}
  \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code Execution.}
\end{figure}

%A opção  define que o match só será feito em pacotes que fazem parte
%de uma conexão já estabelecida entre um cliente e um servidor. Além
%disso, define também que os pacotes devem estar direcionados ao servidor, não ao cliente.

Figure~\ref{fig:rule-jboss} shows a Suricata rule that captures a
JBoss RCE attack.  The option \CodeIn{flow: established, to\_server;}
indicates that a rule match will only occur on packets that (1) flow
on an already-established connection and that (2) flow to the server
as opposed to the client. For the content options the modifiers
\CodeIn{nocase} and \CodeIn{http\_uri} indicate, respectively, that
the matching not be case sensitive and that the string should be
located at the packet field ``Request URI''. \Mar{Lucas, ``nocase''
  tambem eh considerado modificador? Isto nao eh confuso? Na secao
  4.3.4 consideramos modificadores como sinonimo para qualificador de
  campo, nao?}

\tname{} reproduces the malicious JBoss attack using the traffic from
a Nikto script~\cite{nikto}. However, similar JBoss attack is also
available from the Metasploit database~\cite{metasploit}. As in the
other examples, \tname{} uses, as additional inputs, benign traffic
from \Fix{...} and a set of existing Suricata rules from
\Fix{...}. \tname{} initially produces one seed rule containing
\Fix{xx} options, by analyzing the malicious traffic alone. Then, it
uses that seed rule to derive \Fix{yy} rules, each rule containing
subsets of the option in the initial rule. Finally, it ranks that
list. The list below shows the top-5 rules that \tname{}
produces.\Mar{<- check/revise}

\Fix{list rules}

Note that the golden rule is in position \Fix{...} on this
case. \Mar{explain syntactical difference, if exist}
